SAT Solving for Vesicle Traffic Systems in Cells

### 2017-06-21 paper/encoding.tex
added text

\begin{df}
  A VTS $G$ is tuple $(\nodes,\mols,\edges,\nlabel,\pair,\edgef,\nodef)$, where
  \begin{itemize}
  \item $\nodes$ is a set of nodes representing compartments in the VTS,
  \item $\mols$ is the set of molecules flowing in the network, 
  \item $\nlabel : \nodes \maps \powerset{\mols}$ defines the molecules present in the nodes  
  \item $ \nodes \times (\powerset{\mols}-\emptyset) \times \nodes \subseteq \edges$ is the set of edges
    with molecule sets as labels
  \item $\pair : \mols \times \mols$ is pairing relating
  \item $\nodef : \mols \maps \powerset{\mols} \maps \booleans$ is activity maps for nodes
  \item $\edgef : \mols \maps \powerset{\mols} \maps \booleans $ is activity maps for edges
  \end{itemize}
\end{df}
$\pair$ defines which molecules can fuse with which molecules. 
$\nodef$ and $\edgef$ are used to define activity of molecules on nodes and edges.
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and $\nodef(k,\nlabel(n))$ is true.
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and $\edgef(k,M')$ is true.

@@

A {\em path} in $G$ is a sequence $n_1,...,n_\ell$ of nodes 
such that $(n_i,\_,n_{i+1}) \in \edges$ for each $ 0 < i < \ell$.
%
For a molecule $m \in M$,
An {\em $m$-path} in $G$ is a sequence $n_1,...,n_\ell$ of nodes 
such that $(n_i,M',n_{i+1}) \in \edges$ and $m \in M'$ for
each $ 0 < i < \ell$.
%
A node $n'$ is {\em reachable} from node $n$ in $G$ if there is a path
$n,...,n'$ in $G$.
%
A node $n'$ is {\em $m$-reachable} from node $n$ in $G$ if there is a
$m$-path $n,...,n'$ in $G$.
%
We call $G$ {\em stable} if for each $(n,M',n') \in \edges$ and $m \in M'$,
$n$ is $m$-reachable from $n'$.
%
We call $G$ {\em connected} if for each $n,n' \in \nodes$,
$n'$ is reachable from $n$ in $G$.
%
We call $G$ $k$-connected if for each $E' \subseteq \edges$ and $|E'| < k$,
$(\nodes,\mols,\edges-E',\nlabel,\pair,\edgef,\nodef)$ is connected.
%
An edge $(n,M',n') \in \edges$ {\em fuses} with its destination node $n'$
if

State the SAT problem

\subsection{Boolean satisfiability of the problem}

@@

\subsection{Network description}

>>>>

\begin{df}
  A VTS $G$ is tuple $(\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$, where
  \begin{itemize}
  \item $\nodes$ is a set of nodes representing compartments in the VTS,
  \item $\mols$ is the set of molecules flowing in the network, 
  \item $\nlabel : \nodes \maps \powerset{\mols}$ defines the molecules present in the nodes  
  \item $ \nodes \times (\powerset{\mols}-\emptyset) \times \nodes \subseteq \edges$ is the set of edges
    with molecule sets as labels
  \item $\pairs : \mols \times \mols$ is pairing relating
  \item $\nodef : \mols \maps \powerset{\mols} \maps \booleans$ is activity maps for nodes
  \item $\edgef : \mols \maps \powerset{\mols} \maps \booleans $ is activity maps for edges
  \end{itemize}
\end{df}
$\pairs$ defines which molecules can fuse with which molecules.
%
Let $\pairs(M')$ denote $\{m|(m,m') \in P \text{ and } m' \in M'\}$.
%
$\nodef$ and $\edgef$ are used to define activity of molecules on
nodes and edges.
%
A molecule $k$ is {\em active} at node $n$ if $k \in \nlabel(n)$ and
$\nodef(k,\nlabel(n))$ is true.
%
A molecule $k$ is {\em active} at edges $(n,M',n')$ if $k \in M'$ and
$\edgef(k,M')$ is true.

@@

A {\em path} in $G$ is a sequence $n_1,...,n_\ell$ of nodes 
such that $(n_i,\_,n_{i+1}) \in \edges$ for each $ 0 < i < \ell$.
%
For a molecule $m \in M$,
An {\em $m$-path} in $G$ is a sequence $n_1,...,n_\ell$ of nodes 
such that $(n_i,M',n_{i+1}) \in \edges$ and $m \in M'$ for
each $ 0 < i < \ell$.
%
A node $n'$ is {\em reachable} from node $n$ in $G$ if there is a path
$n,...,n'$ in $G$.
%
A node $n'$ is {\em $m$-reachable} from node $n$ in $G$ if there is a
$m$-path $n,...,n'$ in $G$.
%
We call $G$ {\em stable} if for each $(n,M',n') \in \edges$ and $m \in M'$,
$n$ is $m$-reachable from $n'$.
%
An edge $(n,M',n') \in \edges$ {\em fuses} with its destination node $n'$
if there are molecules $m,m' \in \mols$ such that $m$ is active in
$(n,M',n')$, $m'$ is active in $n'$, and $(m,m') \in \pairs$.
%
We call $G$ {\em well-fused} if each edge $(n,M',n') \in \edges$ fuses
with non-empty fusing molecules $M'' \subseteq M'$
and $\pairs(M'')$ are not active in any other node.

%
We call $G$ {\em connected} if for each $n,n' \in \nodes$,
$n'$ is reachable from $n$ in $G$.
%
We call $G$ $k$-connected if for each $E' \subseteq \edges$ and $|E'| < k$,
VTS $(\nodes,\mols,\edges-E',\nlabel,\pairs,\edgef,\nodef)$ is connected.

\paragraph{Search problem}

Based upon earlier discussion, we need to answer the following search
question for ......\ashu{match with the earlier text}
For a given $k$, size $\nu$, and molecule number $\mu$,
we are searching for stable and well-fused VTS
$G = (\nodes,\mols,\edges,\nlabel,\pairs,\edgef,\nodef)$ such that
$|\nodes| = \nu$, $|\mols = \mu|$, and
$G$ is not $k$-connected.

\subsection{Boolean satisfiability of the search problem}

We translate the search problem into a Boolean satisfiability
problem and use SAT solvers to find the satisfying VTSs.
%
We will first present the variables used to encode the
VTSs and the properties.
%
Afterwards, we will present the formulas corresponding to the
properties.

\subsubsection{Variables for VTS description}
%
We assume that size of the graph is $\nu$ and number of molecules is
$\mu$.
%
Furthermore, we also limit the maximum number $\pi$ of edges present
between two nodes.
%
Here, we list the vector of Boolean variables and uninterpreted function symbols
that encode the VTSs.
\begin{enumerate}

\item $n_{i,m}$ indicates if $m \in \nlabel(i)$
\item $e_{i,j,q}$ indicates if $q$th edge exists between $i$ and $j$.
\item $e_{i,j,q,m}$ indicates if $q$th edge between $i$ and $j$ contains molecule $m$.

\item $p_{k,k'}$ indicates if $(m,m') \in \pairs$
\item uninterpreted Boolean functions $\nodef_m : \booleans^\mu \maps \booleans$
encoding $\nodef(m)$ map
\item uninterpreted Boolean functions $\edgef_m : \booleans^\mu \maps \booleans$
encoding $\edgef(m)$ map
\end{enumerate}

We also have auxiliary variables that will help us encode the properties 
\begin{enumerate}
\item dump = d{i,j,q}
      to represent drop of edge{i,j,q} from the graph we will use d{i,j,q}.

\item active\_node = a{i,k} 
      kth molecule is active on the node i.

\item presence\_edge = e{i,j,q,k}
      There is an qth edge between node i and j with kth molecule present.

\item active\_edge = b{i,j,q,k}
      There is an qth edge between node i and j with kth molecule active.

\item reachability = r{i,j,k,z}
      Node i and j are reachable with kth molecule present in z steps.

@@
